---
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Probability and Statistics

# Lab Assignment 1: Naive Bayes Classifier

### *Anastasiia Shvets, Anastasiia Beheni, Nazarii Peniaha*

## Introduction

During the past three weeks, we learned a couple of essential notions
and theorems, and one of the most important among them is the *Bayes
theorem*.

One of its applications is **Naive Bayes classifier**, which is a
probabilistic classifier whose aim is to determine which class some
observation probably belongs to by using the Bayes formula:
$$\mathsf{P}(\mathrm{class}\mid \mathrm{observation})=\frac{\mathsf{P}(\mathrm{observation}\mid\mathrm{class})\mathsf{P}(\mathrm{class})}{\mathsf{P}(\mathrm{observation})}$$

Under the strong independence assumption, one can calculate
$\mathsf{P}(\mathrm{observation} \mid \mathrm{class})$ as
$$\mathsf{P}(\mathrm{observation}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i), \qquad \mathsf{P}(\mathrm{observation} \mid \mathrm{class}) = \prod_{i=1}^{n} \mathsf{P}(\mathrm{feature}_i \mid \mathrm{class}),$$
where $n$ is the total number of features describing a given
observation. Thus, $\mathsf{P}(\mathrm{class}|\mathrm{observation})$ now
can be calculated as

$$\mathsf{P}(\mathrm{class} \mid \mathrm{\mathrm{observation}}) = \mathsf{P}(\mathrm{class})\times \prod_{i=1}^{n}\frac{\mathsf{P}(\mathrm{feature}_i\mid \mathrm{class})}{\mathsf{P}(\mathrm{feature}_i)}\tag{1}$$

## Data description

-   **0 - authors** This data set consists of citations of three famous
    writers: Edgar Alan Poe, Mary Wollstonecraft Shelley and HP
    Lovecraft. The task with this data set is to classify a piece of
    text with the author who was more likely to write it.


```{r}
# here goes a list of recommended libraries,
# though you may install other ones if they are needed
library(tidytext)
library(readr)
library(dplyr)
library(ggplot2)
```

```{r}
# list.files(getwd())
list.files("data/0-authors")
```

```{r}
test_path <- "data/0-authors/test.csv"
train_path <- "data/0-authors/train.csv"

stop_words <- read_file("stop_words.txt")
# https://stackoverflow.com/questions/27195912/why-does-strsplit-return-a-list
splitted_stop_words <- strsplit(stop_words, split='\n')
splitted_stop_words <- splitted_stop_words[[1]]
```

```{r}
train <-  read.csv(file = train_path, stringsAsFactors = FALSE)
test <-  read.csv(file = test_path, stringsAsFactors = FALSE)
```

```{r}
# note the power functional features of R bring us! 
tidy_text <- unnest_tokens(train, 'splitted', 'text', token="words") %>%
             filter(!splitted %in% stop_words)

counter <- tidy_text %>% count(splitted,sort=TRUE)

#some other data pre-processing

lovecraft <- subset(tidy_text, author == "HP Lovecraft") %>% count(splitted, sort = TRUE)
poe <- subset(tidy_text, author == "Edgar Alan Poe") %>% count(splitted, sort = TRUE)
mary <- subset(tidy_text, author == "Mary Wollstonecraft Shelley ") %>% count(splitted, sort = TRUE)

counter <- merge(counter, lovecraft, by = "splitted", all = TRUE)
counter <- merge(counter, poe, by = "splitted", all = TRUE)
counter <- merge(counter, mary, by = "splitted", all = TRUE)

colnames(counter)[1] = "word"
colnames(counter)[2] = "total"
colnames(counter)[3] = "Lovecraft"
colnames(counter)[4] = "Poe"
colnames(counter)[5] = "Shelley"
counter[is.na(counter)]<-0


```

## Classifier implementation

```{r}
naiveBayes <- setRefClass("naiveBayes",
       fields = list(),
       methods = list(
                    fit = function(X, y){
                      total_author_amount_vector <<- c(sum(counter$Lovecraft), sum(counter$Poe), sum(counter$Shelley))
                      total_amount_of_unique_words <<- nrow(counter)
                      total_amount_of_unique_words
                      
                      counter$Prob_word_given_Lovecraft <<- with(counter, (Lovecraft+1)/(total_author_amount_vector[1]+total_amount_of_unique_words))
                      counter$Prob_word_given_Poe <<- with(counter, (Poe+1)/(total_author_amount_vector[2]+total_amount_of_unique_words))
                      counter$Prob__word_given_Shelley <<- with(counter, (Shelley+1)/(total_author_amount_vector[3]+total_amount_of_unique_words))
            
                      total_amount_of_words <<- sum(counter$total)
          
                      prob_author_vector <<- c(total_author_amount_vector/total_amount_of_words)
                      },
                    
                    predict = function(message){
                      message <- strsplit(message, " ")[[1]]
                      message_data <- data_frame(text = message)
                      new_splitted_mes <- unnest_tokens(message_data, 'text', 'text', token="words") %>% 
                                      filter(!text %in% splitted_stop_words)
                      
                      prob_message_lovecraft = 1
                      prob_message_poe = 1
                      prob_message_shelley= 1
            
                      for (wrd in new_splitted_mes$text) {
                        
                          word_prob_lovecraft <- ifelse(wrd %in% counter$word,
                            counter %>% filter(word==wrd) %>% pull(Prob_word_given_Lovecraft),
                            1/(total_author_amount_vector[1]+total_amount_of_unique_words))
                          
                          word_prob_poe <- ifelse(wrd %in% counter$word,
                            counter %>% filter(word==wrd) %>% pull(Prob_word_given_Poe),
                            1/(total_author_amount_vector[2]+total_amount_of_unique_words))
                          
                          word_prob_shelley <- ifelse(wrd %in% counter$word,
                            counter %>% filter(word==wrd) %>% pull(Prob_word_given_Shelley),
                            1/(total_author_amount_vector[3]+total_amount_of_unique_words))
                          
                        
                      prob_message_lovecraft = prob_message_lovecraft * word_prob_lovecraft
                      prob_message_poe = prob_message_poe * word_prob_poe
                      prob_message_shelley= prob_message_shelley * word_prob_shelley
                      
                    }
                      
                      lst <- list(prob_message_lovecraft*prob_author_vector[1], prob_message_poe*prob_author_vector[2],
                                prob_message_shelley*prob_author_vector[3])
                      index <- which.max(lst)
                      lst_authors <- list("HP Lovecraft", "Edgar Alan Poe", "Mary Wollstonecraft Shelley ")
                      return(lst_authors[index])
                    
            }
                    
                    score = function(X_test, y_test){
                      path = "data/0-authors/test.csv"
                      data_from_file <- read.csv(path)
                      data_from_file <- data_from_file[, (3:4)]
                      for (i in 1:nrow(data_from_file)) {
                        data_from_file[i, "predicted"] = naiveBayes()$predict(data_from_file$text[i])
                        data_from_file[i, "result"] = ifelse(data_from_file$author[i] == data_from_file$predicted[i], 1, 0)
                      }
                      data_from_file
                      
                      # checker <- read.csv("data/0-authors/final_data_fr.csv")
                      
                      edgar_data <- data_from_file %>% filter(author == "Edgar Alan Poe") %>% select(author, predicted, result)
                      edgar_prob = sum(edgar_data$result)/nrow(edgar_data)
                      
                      mary_data <- data_from_file %>% filter(author == "Mary Wollstonecraft Shelley ") %>% select(author, predicted, result)
                      mary_prob = sum(mary_data$result)/nrow(mary_data)
                      
                      hp_data <- data_from_file %>% filter(author == "HP Lovecraft") %>% select(author, predicted, result)
                      hp_prob = sum(hp_data$result)/nrow(mary_data)
                      
                      author <- c("HP Lovecraft", "Mary Wollstonecraft Shelley ", "Edgar Alan Poe")
                      probability <- c(hp_prob, mary_prob, edgar_prob)
                      
                      new_data <- data.frame(author, probability)
                      new_data %>% 
                        ggplot(aes(author, probability, colour = probability))+
                        geom_bar(stat = "identity", aes(fill = probability))
                      
                }
                    
))

model = naiveBayes()
model$fit()
model$score()
```

## Measure effectiveness of your classifier
```{r}
accuracy_path = "data/0-authors/final_data_fr.csv"
print(getwd())
accur <- read.csv(accuracy_path)
print("Effectiveness of our classifier =")
print(sum(accur$result)/nrow(accur)*100)
```

## Conclusions
- We used Naive Bayes Classifier which uses Bayes's formula for conditional probability to determine the authorship of the given text excerpt
- Pros of this approach:
    The algoright is easy to understand and implement
    We got pretty high accuracy (83%)
    It can be used for both binary and multi-class classifications
- Cons:
    It assumes that all the features are independent. While it might sound great in theory, in real life, you’ll hardly find a set of independent features
    We have to deal (by Laplace smoothing) with the problem when the word wasn't presented in train set so our probability equals to one

